### 2013-11-21 descriptive_analysis.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attributions")
library(knitr)
knit2html('descriptive_analysis.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))

knitpandoc("descriptive_analysis",toc)
```

>>>>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attributions")
library(knitr)
library(markdown)
knit2html('descriptive_analysis.Rmd', 
          stylesheet="~/workspace/web/css/rmarkdown.css",
          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
```

### 2013-11-21 finalPlots.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attributions/")
library(knitr)
knit2html('finalPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))

>>>>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attributions/")
library(knitr)
library(markdown)
knit2html('finalPlots.Rmd', 
          stylesheet="~/workspace/web/css/rmarkdown.css",
          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
```

### 2013-11-21 loadClean.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

@@

### ------ BEGIN 2006 ---------- ####
# $sex
lits1$sex <- lits1$q102_1
# $income
library(plyr)
medianData <- ddply(lits1,~cntry,summarise,incMedian=median(q211, na.rm=TRUE))
lits1 <- merge(lits1,medianData,by="cntry",all.x=TRUE)

lits1$income2[lits1$q211 < lits1$incMedian]  <- 'Low'
lits1$income2[lits1$q211 >= lits1$incMedian]  <- 'High'

@@

lits2$income2[lits2$q227 >= 1 & lits2$q227 <= 5]  <- 'Low'
lits2$income2[lits2$q227 >= 6 & lits2$q227 <= 10]  <- 'High'

@@

### ------ BEGIN 2006 ---------- ####
## perGini
library(laeken)
perGini2006 <- as.data.frame(gini(inc="q211",
                      weights = "weight_1",
                      data = lits1,
                      na.rm = TRUE,
                      breakdown="cntry")[2])

names(perGini2006) <- c("cntry","perGini")
perGini2006

library(laeken)
perGini2010 <- as.data.frame(gini(inc="q227_rec",
                      weights = "weight",
                      data = lits2,
                      na.rm = TRUE,
                      breakdown="cntry")[2])

names(perGini2010) <- c("cntry","perGini")
perGini2010

@@

lits1$incsour3[lits1$q210b %in% c('state provided social benefits',
                                  'unemployment benefits',
                                  'pensions',
                                  'community/privately provided social benefits',
                                  'help from charities and NGOs',
                                  'help from relatives/friend in the country',
                                  'help from relatives/friends abroad')] <- "dependent"
lits1$incsour3[!(lits1$q210b %in% c('state provided social benefits',
                                  'unemployment benefits',
                                  'pensions',
                                  'community/privately provided social benefits',
                                  'help from charities and NGOs',
                                  'help from relatives/friend in the country',
                                  'help from relatives/friends abroad'))] <- "notDependent"

@@

lits1$edu2[lits1$q501 %in% c('compulsory school education',
                         'no degree/education')] <- 'no or compulsory'
lits1$edu2[!(lits1$q501 %in% c('compulsory school education',
                         'no degree/education'))] <- 'higher'

@@

lits2$incsour3 <- factor(lits2$incsour3, 
                         levels=c("dependent",
                                  "notDependent"))
summary(lits2$incsour3)
###

@@

lits2$edu2[lits2$q515 == 1 | lits2$q515 == 2]  <- "no or compulsory"
lits2$edu2[lits2$q515 >= 3]  <- "higher"

@@

```{rLoadClean5, eval=TRUE}
## Data for the analysis
df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
                             "sex","age",
                             "future.diff","past.diff",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             "income2",
                             "crise",
                             "poverty"
                             ))
# filter out not wanted countries
df <- df[!is.na(df$group_analysis),] # valitaan vaan group1 -muuttujan maat
df$cntry <- factor(df$cntry)
summary(df)

@@

```{rLoadClean6, eval=TRUE}
###
library(reshape2)
library(survey)
d.df <- svydesign(id = ~SerialID, 
                  weights = ~weight, 
                  data = df)

summary309 <- data.frame(prop.table(svytable(~q309 + 
                                               cntry, d.df), 
                                    2) * 100)

summary309 <- summary309[!is.na(summary309$Freq), ]
summary309w <- dcast(summary309, 
                     cntry ~ q309, 
                     value.var="Freq")
names(summary309w) <- c("cntry","socialBlame",
                        "individualBlame","individualFate",
                        "socialFate","not.stated","dont.know")

summaryPoverty <- data.frame(prop.table(svytable(~poverty + 
                                                   cntry, d.df), 
                                        2) * 100)
summaryPoverty <- summaryPoverty[!is.na(summaryPoverty$Freq), ]

summaryPovertyw <- dcast(summaryPoverty, cntry ~ poverty, value.var="Freq")
names(summaryPovertyw) <- c("cntry","socialBlame","individualBlame","individualFate","socialFate")

@@

```{rLoadClean7, eval=TRUE}
#######
# Quality of governance
library(rustfare)
datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
                   'Kazakhstan','Tajikistan',
                   'Uzbekistan','Estonia','Latvia','Lithuania',
                   'Poland','Czech Republic','Slovakia','Hungary',
                   'Slovenia','Romania','Bulgaria',
                   'Italy','Germany','United Kingdom',
                   'Sweden','France'), 
       indicator = c("ffp_fsi", # Failed States Index
                     "wbgi_vae", # World Bank - Voice and Accountability
                     "undp_hdi", # UNDP Human Development Index
                     "fh_polity2", # Democracy (Freedom House/Polity)
                     "uw_gini"))
# subset the latest value per cntry per indicator
library(plyr)
datQogSum <- ddply(datQog, c("cname","indicator"), 
                   subset, year == max(year))

datQogSum.w <- dcast(datQogSum, 
                     cname ~ indicator, 
                     value.var="value")

@@

datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
                                    levels=c("Western Europe",
                                             "CEE",
                                             "CIS"))

df.qog <- datQogSum.w

names(df.qog)[1] <- "cntry"

@@

```{rLoadClean8, eval=TRUE}
## GDP change from WB
### We need iso2c countrycode
library(countrycode)
cntry <- levels(df$cntry) # countrynames
iso2c <- countrycode(cntry, "country.name", "iso2c")
iso2c # copy paste into WDI search
library(WDI)
# WDIsearch(string='gdp', field='name', cache=NULL)
wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
          country = c("AL","AM","AZ","BY","BA","BG","HR",
                      "CZ","EE","FR","GE","DE","GB","HU",
                      "IT","KZ","KG","LV","LT","MK","MD",
                      "MN","ME","PL","RO","RU","RS","SK",
                      "SI","SE","TJ","TR","UA","UZ"), 
           start = 2007, end = 2010)

library(reshape2)
names(wb) <- c("iso2c","cntry","var","year")
wb.w <- dcast(wb, 
              cntry ~ year, 
              value.var="var")

names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
gdp <- wb.w[,c(1,6)]

gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'

@@

```{rLoadClean9}

# yhdistetään nää aikaisempien datojen kanssa
macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
macro <- merge(macro,df.qog,by="cntry")
macro <- merge(macro,perGini2010,by="cntry")

df <- merge(df,macro,by="cntry")

rm(lits2)
rm(df.qog)
rm(df.cross.postsoc)
rm(gdp)
rm(t3)
rm(summary309)
rm(summary309_rec)
rm(d.df)
rm(wb)
rm(wb.w)
rm(cntry)
rm(iso2c)
rm(lits1)
rm(medianData)
rm(perGini2006)
rm(perGini2010)

>>>>

@@

### ------ BEGIN 2006 ---------- ####
# $sex
lits1$sex <- lits1$q102_1
# $income
library(plyr)
lits1$q211_rec <- lits1$q211
lits1$q211_rec[lits1$q211_rec >= 11] <- NA

medianData <- ddply(lits1,~cntry,summarise,incMedian=median(q211_rec, na.rm=TRUE))
lits1 <- merge(lits1,medianData,by="cntry",all.x=TRUE)

lits1$income2[lits1$q211_rec < lits1$incMedian]  <- 'Low'
lits1$income2[lits1$q211_rec >= lits1$incMedian]  <- 'High'

@@

@@

@@

lits1$incsour3[lits1$q210b %in% c('state provided social benefits',
                                  'unemployment benefits',
                                  'community/privately provided social benefits',
                                  'help from charities and NGOs',
                                  'help from relatives/friend in the country',
                                  'help from relatives/friends abroad')] <- "dependent"
lits1$incsour3[!(lits1$q210b %in% c('state provided social benefits',
                                  'unemployment benefits',
                                  'community/privately provided social benefits',
                                  'help from charities and NGOs',
                                  'help from relatives/friend in the country',
                                  'help from relatives/friends abroad'))] <- "notDependent"

@@

lits2$incsour3 <- factor(lits2$incsour3, 
                         levels=c("dependent",
                                  "notDependent"))
summary(lits2$incsour3)

# $edu
###
summary(lits1$q501)

lits1$edu2[lits1$q501 %in% c(#'compulsory school education',
                         'no degree/education')] <- 'no or compulsory'
lits1$edu2[!(lits1$q501 %in% c(#'compulsory school education',
                         'no degree/education'))] <- 'higher'

@@

@@

lits2$edu2[lits2$q515 == 1]  <- "no or compulsory"
lits2$edu2[lits2$q515 >= 2]  <- "higher"

@@

```{rLoadClean5, eval=TRUE}
## Data for the analysis

df <- subset(lits2, select=c("SerialID","weight",
                             "cntry","q309",
                             "sex","age",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             "income2",
                             "poverty",
                             "future.diff","past.diff"
                             ))
names(lits1)[2] <- "SerialID"
names(lits1)[1165] <- "weight"

df2 <- lits1[,c("SerialID","weight",
                             "cntry","q309",
                             "sex","age",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             "income2",
                             "poverty"
                             )]

# filter out not wanted countries
df <- df[!is.na(df$group_analysis),] # valitaan vaan group1 -muuttujan maat
df$cntry <- factor(df$cntry)
summary(df)

@@

```{rLoadClean6, eval=TRUE}
### Dependent var country level values
## 2010
library(survey)
d.df <- svydesign(id = ~SerialID, weights = ~weight, data = df)
q309_10 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
q309_10 <- q309_10[!is.na(q309_10$Freq), ]
q309_10w <- dcast(q309_10, cntry ~ q309, value.var="Freq")
names(q309_10w) <- c("cntry","socialBlame",
                        "individualBlame","individualFate",
                        "socialFate","notStated","dontKnow")

poverty_10 <- data.frame(prop.table(svytable(~poverty + cntry, d.df), 
                                        2) * 100)
poverty_10 <- poverty_10[!is.na(poverty_10$Freq), ]
poverty_10w <- dcast(poverty_10, cntry ~ poverty, value.var="Freq")
names(poverty_10w) <- c("cntry","socialBlameRec","individualBlameRec",
                        "individualFateRec","socialFateRec")
library(laeken)
perGini2010 <- as.data.frame(gini(inc="q227_rec",
                      weights = "weight",
                      data = lits2,
                      na.rm = TRUE,
                      breakdown="cntry")[2])
names(perGini2010) <- c("cntry","perGini")

macro2010 <- merge(q309_10w,poverty_10w,by="cntry",all.x=TRUE)
macro2010 <- merge(macro2010,perGini2010,by="cntry",all.x=TRUE)
macro2010$year <- 2010
### ------ BEGIN 2006 ---------- ####
library(survey)
df2 <- df2[!is.na(df2$weight),]
d.df <- svydesign(id = ~SerialID, weights = ~weight, data = df2)
q309_06 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
q309_06 <- q309_06[!is.na(q309_06$Freq), ]
q309_06w <- dcast(q309_06, cntry ~ q309, value.var="Freq")
names(q309_06w) <- c("cntry","socialBlame",
                        "individualBlame","individualFate",
                        "socialFate","notStated")
q309_06w$dontKnow <- NA

poverty_06 <- data.frame(prop.table(svytable(~poverty + cntry, d.df), 
                                        2) * 100)
poverty_06 <- poverty_06[!is.na(poverty_06$Freq), ]
poverty_06w <- dcast(poverty_06, cntry ~ poverty, value.var="Freq")
names(poverty_06w) <- c("cntry","socialBlameRec","individualBlameRec",
                        "individualFateRec","socialFateRec")
## perGini
library(laeken)
perGini2006 <- as.data.frame(gini(inc="q211_rec",
                      weights = "weight_1",
                      data = lits1,
                      na.rm = TRUE,
                      breakdown="cntry")[2])

names(perGini2006) <- c("cntry","perGini")
macro2006 <- merge(q309_06w,poverty_06w,by="cntry",all.x=TRUE)
macro2006 <- merge(macro2006,perGini2006,by="cntry",all.x=TRUE)
macro2006$year <- 2006

#macro <- rbind(macro2006,macro2010)
### ------ END 2006 ---------- ####
```

@@

```{rLoadClean7, eval=TRUE}
#######
# Quality of governance
library(rustfare)
datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
                   'Kazakhstan','Tajikistan',
                   'Uzbekistan','Estonia','Latvia','Lithuania',
                   'Poland','Czech Republic','Slovakia','Hungary',
                   'Slovenia','Romania','Bulgaria',
                   'Italy','Germany','United Kingdom',
                   'Sweden','France'), 
       indicator = c("wbgi_vae"#, # World Bank - Voice and Accountability
                     #"uw_gini"
                     ))

# datQog <- datQog[datQog$year >= 2000, ]
# table(factor(datQog$cname),factor(datQog$year),factor(datQog$indicator))

# subset the latest value per cntry per indicator
#library(plyr)
#datQogSum <- ddply(datQog, c("cname","indicator"), 
#                   subset, year == max(year))
datQogSum <- datQog[datQog$year %in% c(2006,2009),]
datQogSum$year[datQogSum$year == 2009] <- 2010
names(datQogSum)[3] <- "wbgi_vae"
datQogSum$indicator <- NULL

# datQogSum.w <- dcast(datQogSum, 
#                      cname+indicator ~ year, 
#                      value.var="value")

datQogSum.w <- datQogSum

@@

datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
                                    levels=c("Western Europe",
                                             "CEE",
                                             "CIS"))
df.qog <- datQogSum.w
names(df.qog)[1] <- "cntry"

macro2006 <- merge(macro2006,df.qog[df.qog$year == 2006,],by=c("cntry","year"))
macro2010 <- merge(macro2010,df.qog[df.qog$year == 2010,],by=c("cntry","year"))

@@

```{rLoadClean8, eval=TRUE}
## GDP change from WB
### We need iso2c countrycode
library(countrycode)
cntry <- levels(df$cntry) # countrynames
iso2c <- countrycode(cntry, "country.name", "iso2c")
iso2c # copy paste into WDI search
library(WDI)
# WDIsearch(string='gdp', field='name', cache=NULL)
wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
          country = c("AL","AM","AZ","BY","BA","BG","HR",
                      "CZ","EE","FR","GE","DE","GB","HU",
                      "IT","KZ","KG","LV","LT","MK","MD",
                      "MN","ME","PL","RO","RU","RS","SK",
                      "SI","SE","TJ","TR","UA","UZ"), 
           start = 2003, end = 2010)
library(reshape2)
names(wb) <- c("iso2c","cntry","var","year")
wb$year <- paste("x",wb$year,sep="")
wb.w <- dcast(wb, 
              cntry ~ year, 
              value.var="var")

wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
wb.w$gdpchange03.06 <- wb.w$x2006/wb.w$x2003 * 100 - 100

wb.w$cntry[wb.w$cntry == 'Russian Federation'] <- 'Russia'
wb.w$cntry[wb.w$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
wb.w$cntry[wb.w$cntry == 'United Kingdom'] <- 'Great Britain'
wb.w$cntry[wb.w$cntry == 'Slovak Republic'] <- 'Slovakia'

gdp10 <- wb.w[,c(1,10)]
names(gdp10) <- c("cntry","gdpChange")
gdp06 <- wb.w[,c(1,11)]
names(gdp06) <- c("cntry","gdpChange")

macro2006 <- merge(macro2006,gdp06,by="cntry")
macro2010 <- merge(macro2010,gdp10,by="cntry")
```

@@

```{rLoadClean9}
# yhdistetään nää aikaisempien datojen kanssa
macro <- rbind(macro2006,macro2010)

df <- merge(df,macro[macro$year == 2010,],by="cntry")
df2 <- merge(df2,macro[macro$year == 2006,],by="cntry")

rm(list=setdiff(ls(), c("macro","df","df2")))
